Parks Data

### 2019-07-15 02_parks_data.Rmd
Revise open spaces builder
Calculate trails, playgrounds, courts, etc.

## Cleaning
The map of Flushing Meadows in Queens below shows some of the difficulties of this dataset:
  - First, the park is constructed of multiple outlines that limits the size of 
  the total park. This makes Flushing Meadows appear smaller than Central Park, 
  which is a single large park. 
  - Second, the dataset includes tennis parks and baseball fields as separate 
  facilities.
  - Third, parking lots for stadia (in this case Citi Field) are included as 
  open park space. We should remove these as they are categorically unlike other
  green spaces.
  - Fourth, important green spaces like cemeteries are in a different dataset.

@@

First, we manually remove the park facilities corresponding to stadia and their
associated parking lots, and then filter out facilities that are not explicitly
park boundaries, as identified by the code `4980`.

```{r notparks}
notparks <- c(9498000833, 14498000872, 14498000794, # citi field, Yankee stadium, garage
              9498000631)  # USTA Billie Jean Tennis Center
parks <- parks %>% 
  filter(feat_code == 4980) %>%
  filter(!(source_id %in% notparks))
```

```{r fm2}
leaflet(st_buffer(parks %>% filter(grepl("Flushing Meadows", park_name)), dist = 0.000001)) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(label = ~as.character(park_name), stroke = 0.5)
```

With those areas removed, we build a 30-foot buffer around each park and 
dissolve interior boundaries. This will make parks that are effectively the same
facility a single polygon feature. We also re-calculate the area of the parks in
acres and remove parks smaller than half an acre.

```{r dissolve}
dissolved <- parks %>%
  #filter(grepl("Flushing Meadows", park_name)) %>%
  st_transform(3628) %>%
  st_buffer(30) %>%
  group_by(parknum) %>%
  summarise(area = sum(shape_area),
            park_name = park_name[1]) %>%
  mutate(sqft = st_area(.), acres = units::set_units(sqft, acres)) 
```

```{r fm3}
leaflet(st_transform(dissolved, 4326) ) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(stroke = 0.5)
```

@@

Now we can join the cemeteries data to the parks data and write out the results
```{r write}
rbind(
  dissolved %>% transmute(name = park_name, acres, type = "park"),
  dissolved_cemeteries %>% transmute(name, acres, type = "cemetery")
) %>%
  filter(as.numeric(acres) > 0.5) %>%
  st_simplify(preserveTopology = TRUE) %>%
  write_rds("data/open_spaces.rds")
```

>>>>

The New York City open data portal also contains a list of designated trails 
within parks.

```{r trails}
trails <- read_csv("https://data.cityofnewyork.us/api/views/vjbm-hsyr/rows.csv?accessType=DOWNLOAD")
trails2 <- jsonlite::fromJSON("https://www.nycgovparks.org/bigapps/DPR_Hiking_001.json")
parks_with_trails <- c(trails$ParkID, trails2$Prop_ID)
```

## Cleaning
The map of Flushing Meadows in Queens below shows some of the difficulties of this dataset:
  - First, the park is constructed of multiple outlines that limits the size of 
  the total park. This makes Flushing Meadows appear smaller than Central Park, 
  which is a single large park. 
  - Second, the dataset includes tennis parks and baseball fields as separate 
  facilities. We want to include these as attributes of their containing park
  rather than independent facilities.
  - Third, parking lots for stadia (in this case Citi Field) are included as 
  open park space. We should remove these as they are categorically unlike other
  green spaces.
  - Fourth, important green spaces like cemeteries are in a different dataset.

@@

First, we manually remove the park facilities corresponding to stadia and their
associated parking lots, and then tag facilities that are not explicitly
park boundaries.

```{r notparks}
notparks <- c(9498000833, 14498000872, 14498000794, # citi field, Yankee stadium, garage
              9498000631)  # USTA Billie Jean Tennis Center

parks <- parks %>% 
  filter(!(source_id %in% notparks)) %>%
  mutate(
    courts = ifelse(grepl("Courts", landuse), TRUE, FALSE),
    playgrounds = ifelse(grepl("Playground", landuse), TRUE, FALSE),
    trails = ifelse(parknum %in% parks_with_trails, TRUE, FALSE)
  ) 
```

With those areas removed, we build a 30-foot buffer around each park and 
dissolve interior boundaries based on the park number. Tennis courts, etc. become
boolean variables indicating whether the park has the facilities or not. This
will make parks that are effectively the same facility a single polygon feature.
We also re-calculate the area of the parks in acres and remove parks smaller
than half an acre.

```{r dissolve}
dissolved <- parks %>%
  filter(!is.na(parknum)) %>%
  group_by(parknum) %>%
  filter(shape_area > 0.5 * 43560 | courts) %>%
  summarise(
    area = sum(shape_area), 
    courts = any(courts), 
    playgrounds = any(playgrounds),
    trails = any(trails),
    park_name = park_name[1]
  ) %>%
  mutate(sqft = st_area(.), acres = units::set_units(sqft, acres)) %>%
  st_transform(3628)
```

```{r fm3}
leaflet(st_transform(dissolved, 4326) ) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(stroke = 0.5,)
```

@@

Now we can join the cemeteries data to the parks data and write out the results
```{r write}
rbind(
  dissolved %>% 
    transmute(id = parknum, name = park_name, acres, courts, playgrounds, trails, 
              type = "park"),
  dissolved_cemeteries %>% 
    transmute(id = source_id, name, acres, courts = FALSE, playgrounds = FALSE, 
              trails = FALSE, type = "cemetery")
  ) %>%
  filter(as.numeric(acres) > 0.5) %>%
  st_simplify(preserveTopology = TRUE) %>%
  write_rds("data/open_spaces.rds") %>%
  geojsonio::geojson_write(file = "data/open_spaces.geojson")
```
